### 2021-07-30 response.tex

\rev{Reviewer \#1}

\revtext{This is a review of ``Roles of generation-interval distributions in shaping relative epidemic strength, speed, and control of new SARS-CoV-2 variants'' by Park and others for Royal Society Interface.  This is the latest in an ongoing series of papers (Champredon 2015, Champredon 2018, Park 2019, Dushoff 2021, ...) in which the last author has considered aspects of incubation, growth, and reproduction in SIR epidemic theory.  In this manuscript, the authors use theory to anticipate the effects of COVID-19 variants where the distribution of infectivity evolves.}

\revtext{The manuscript is excellently written. The figures are clear.  All the mathematics seems satisfactory.  Unfortunately, the manuscript does not make its case for novel results.}

Thank you for the review.
We have tried to frame our results better, and to clarify both the novelty and importance of this study.
The primary purpose of our study is two-fold: (1) to illustrate the impact of uncertainty in the generation interval in shaping the population-level transmission advantage of new variants and (2) to evaluate when it is appropriate to hold relative strength or speed constant.
While several studies of new variants have assumed constant relative speed (e.g., by fitting a logistic growth model), we argue that holding relative strength fixed is generally more appropriate given that the dynamics of SARS-CoV-2 have been driven primarily by strength-like interventions.
We also show that this relationship changes under speed-like interventions---that is, relative strength changes but relative speed remains constant.
We believe that the addition of a new simulations (Figure 5) and the revised text better highlight the novelty of our work.

@@

\revtext{Other comments:}

\revtext{Topic on which there is a considerable confusion in terminology in the literature.  In addition to "generation interval", terms "generation time", "transmission interval", "serial interval", and others have been used in the literature, all with related by different definitions.  This manuscripts definition "time between infection and transmission" implies a single transmission event, but Kenah pointed out that time-contraction appears when allowing multiple transmission events (see Champredon 2015).  A clearer definition up front will help readers.  Fortunately, the results don't depend on the specific definition.}

Thank you for the suggestions. We have tried to clarify what we mean by the generation interval. In particular, often-mentioned time contractions of generation intervals apply to the forward realized generation interval, which measure time between actual infection events, rather than the intrinsic generation interval, which measure the intrinsic infectiousness of infected individuals. Under strength-like interventions, the intrinsic generation interval does not vary across time, even though the realized generation interval contracts. These distinctions are explained in Champredon 2015, and we have tried to clarify these ideas again here in Section 2:

@@

We further note that a fixed generation-interval distribution does not rely on the assumption that the rate of depletion of susceptibles over time is negligible.
This is due to an ongoing confusion between the intrinsic and realized generation interval as we explained above.
For example, as shown in Champredon et al 2015 and 2018, the standard SEIR model can be equivalently written in terms of the renewal equation using a fixed \emph{intrinsic} generation-interval distribution, and we can still observe the contraction of the \emph{realized} generation interval for this model.
We have added a paragraph in Section 2 to clarify this distinction (please see response to reviewer 1 for text).

\revtext{(page 2, lines 55–60) It would help to clarify what is meant by “strengthlike” and “speed-like” in interventions. From the examples, it appears that
speed-like interventions work by identifying potential infections quickly
(effectively ending the infectious period early while not necessarily altering
the hazard of infectious contact) and strength-like interventions work by
reducing the hazard of infectious contact (while not necessarily altering
the duration of infectiousness). A good explanation of these ideas is on
lines 228–231 on page 10.}

We now define them more clearly throughout the text, including in the introduction:

``For example, when epidemic strength $\RR$ is fixed, assuming longer generation intervals leads to a slower epidemic growth (lower $r$), making the epidemic look easier to control with a constant-speed intervention, which isolates infected individuals at a constant hazard and therefore affects late transmission disproportionately---as shown in \cite{doi:10.1098/rspb.2020.1556}, the hazard of isolation must be greater than the epidemic speed $r$ to reach the $r=0$ threshold.
In this same case, however, the uncertainty in the generation-interval distribution is expected to have no effects on the effectiveness of a constant-strength intervention, which reduces the hazard of infectious contacts and therefore affects transmission by an equal amount regardless of age of infection.
Conversely, when epidemic speed $r$ is fixed, assuming longer generation intervals leads to a higher estimate of epidemic strength ($\RR$), making the epidemic look harder to control with a constant-strength interventions (but not with a constant-speed interventions, whose effectiveness is fixed if speed is fixed).
Constant strength and speed are idealized representations of real-life interventions, which can be either strength-like (e.g., vaccination and social distancing) or speed-like (e.g., contact tracing and isolation) depending on how their effectiveness varies across the generation interval (see \cite{doi:10.1098/rspb.2020.1556} and Discussion).''
